---
title: "FA5 Probability"
author: "Lindsay Faith Bazar"
date: "March 03 2024"
output: 
  cleanrmd::html_document_clean:
    theme: kacit
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}

percentoferrors <- c(1, 2, 1.5)  
percentofmessages <- c(40, 25, 35) 


table <- data.frame(Server = c("Server 1", "Server 2", "Server 3"),
                 `Percentage of Messages` = percentofmessages,
                 `Percentage of Errors` = percentoferrors)

```

**6) An email message can travel through one of three server routes. The percent- age of errors in each of the servers and the percentage of messages that travel through each route are shown in the following table. **

```{r}

table

```

**Assume that the servers are independent.**

*(a) Probability of receiving an email containing an error*

Let's first convert the percentage to decimal by dividing the probability of the server and the message by a hundred. 

Next, we will designate `proboferrors` and `probofmessages`, respectively, for the probability of errors and messages. The message probability includes the message probability for each server, and the error probability includes the error probabilities for each server.

To calculate for the overall probability of receiving an email containing an error denoted as `proboferror`, we will take the sum of the product of proboferrors and probofmessages. 

```{r}
proboferrors <- percentoferrors / 100
probofmessages <- percentofmessages / 100

proboferror <- sum(proboferrors * probofmessages)

proboferror * 100
```

The probability of receiving an email containing an error is `1.425%`  

*(b) Probability that a message will arrive without error*

The probability of a message with no error is denoted as `probofnoerror`. Since the probability of no error is just the complement of receiving a message containing an error, it is therefore calculated as 1 minus the `proboferror`.

```{r}
probofnoerror <- 1 - proboferror

probofnoerror * 100 

```

The probability that a message will arrive without error is `98.575%`

*(c) Probability that if a message arrives without error, it was sent through server 1*

`probofserver1noerror`calculates the conditional probability that if a message arrives without error, it was sent through server 1. This is done by taking the probability of no error for server 1 `(1 - proboferrors[1])`, multiplying it by the probability of the message being sent through server 1 `(probofmessages[1])`, and dividing it by the overall probability of no error `(probofnoerror)`. This gives us the probability of the message being sent through server 1 given that it arrived without error.

```{r}

probofserver1noerror <- (1 - proboferrors[1]) * probofmessages[1] / probofnoerror

probofserver1noerror * 100

```

The probability that if a message arrives without error, it was sent through server 1 is approximately `40.17%`

**9) A software company surveyed managers to determine the probability that they would buy a new graphics package that includes three-dimensional graphics. About 20% of office managers were certain that they would not buy the package, 70% claimed that they would buy, and the others were undecided. Of those who said that they would not buy the package, only 10% said that they were interested in upgrading their computer hardware. Of those interested in buying the graphics package, 40% were also interested in upgrading their computer hardware. Of the undecided, 20% were interested in upgrading their computer hardware.**


*Let A denote the intention of not buying,*

```{r}
A <- 0.20
```

*B the intention of buying,*

```{r}
B <- 0.70
```

*C the undecided,*

```{r}
C <- 0.10
```

*and G the intention of upgrading the computer hardware.*

```{r}
PofGA <- 0.10
PofGB <- 0.40
PofGC <- 0.20
```

*(a) Calculate the probability that a manager chosen at random will not upgrade the computer hardware (P(G̅)).*

To calculate the probability that a manager chosen at random will not upgrade the computer hardware (P(G̅)), we need to consider the complement of the event that a manager will upgrade the computer hardware (P(G)), which means we will need to find (P(G)) first.

We can use the law of total probability to calculate P(G) denoted as `PofG`:

```{r}
PofG <- PofGA * A + PofGB * B + PofGC * C

PofG
```

Now that we have solved P(G), we can now find the complement (P(G̅). 

```{r}

PofnotG <- 1 - PofG

PofnotG
```

The probability that a manager chosen at random will not upgrade the computer hardware is `68%`

*(b) Explain what is meant by the posterior probability of B given G, P(B|G). *

The posterior probability of B given G, P(B|G), represents the probability of a manager intending to buy the package given that they are interested in upgrading their computer hardware.

**13) A malicious spyware can infect a computer system though the Internet or through email. The spyware comes through the Internet 70% of the time and 30% of the time, it gets in through email. If it enters via the Internet the anti-virus detector will detect it with probability 0.6, and via email, it is detected with probability 0.8.**

*(a) What is the probability that this spyware infects the system?*

`PofI` represents the probability of a malicious spyware entering through the Internet, `PofE` represents the probability of spyware entering through email, `PDwithI` represents the probability of detection given it entered through the Internet, and `PDwithE` represents the probability of detection given it entered through email.


```{r}

PofI <- 0.7  
PofE <- 0.3  
PDwithI <- 0.6  
PDwithE <- 0.8  

```

To calculate the probability that the spyware infects the system, we use the law of total probability to find the probability of a spyware infecting the system.

```{r}

spyware <- PofI * PDwithI + PofE * PDwithE

spyware * 100
```

The probability that the spyware infects the system is `66%`

*(b) If the spyware is detected, what is the probability that it came through the Internet?*

`PofDI` is the probability of getting detected through the Internet. We use Bayes' theorem to calculate the probability that the spyware came through the Internet given that it's detected.

```{r}

PofDI <- (PDwithI * PofI) / spyware

PofDI * 100

```

The probability of a spyware being detected and it came through the Internet is approximately `63.64%`
