---
title: "FA8 Probability"
author: "Lindsay Faith Bazar"
date: "April 18 2024"
output: 
  cleanrmd::html_document_clean:
    theme: kacit
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. An analogue signal received at a detector, measured in microvolts, is normally distributed with mean of 200 and variance of 256.

Here, we define the mean and variance of the normal distribution. We also calculate the standard deviation by taking the square root of the variance.

```{r}
mean <- 200
variance <- 256
std_dev <- sqrt(variance)
```

Then, we assign the cumulative distribution function for the normal distribution to the variable `cdf`. In R, `pnorm` calculates the cumulative distribution function for a given value, mean, and standard deviation.

```{r}
cdf <- pnorm
```

*(a) What is the probability that the signal will exceed 224 µV?*  

To find this, we use the cumulative distribution function `(cdf)` to find the probability that the signal is less than or equal to 224 µV. Then, we subtract this probability from 1 to find the probability that the signal exceeds 224 µV.

```{r}
probg224 <- 1 - cdf(224, mean, std_dev)
probg224 * 100
```

The probability that the signal will exceed 224 µV is `6.68%`

*(b) What is the probability that it will be between 186 and 224 µV?*

Here, we find the probability that the signal is less than or equal to 224 µV and subtract from it the probability that the signal is less than or equal to 186 µV.

```{r}
prob186_224 <- cdf(224, mean, std_dev) - cdf(186, mean, std_dev)
prob186_224 * 100
```

The probability that the signal will exceed 224 µV is `74.24%`

*(c) What is the micro voltage below which 25% of the signals will be?*

We use the quantile function `qnorm` to find the voltage value below which 25% of the signals fall.

```{r}
microvoltage25 <- qnorm(0.25, mean, std_dev)
microvoltage25
```

The micro voltage below which 25% of the signals will be `189.208 µV.`

*(d) What is the probability that the signal will be less than 240 µV, given that it is larger than 210 µV?*

We calculate this probability by dividing the probability that the signal is between 210 and 240 µV by the probability that the signal is greater than 210 µV.

```{r}
probl240g210 <- (pnorm(240, mean, std_dev) - pnorm(210, mean, std_dev)) / (1 - pnorm(210, mean, std_dev))
probl240g210 * 100
```

The probability that the signal will be less than 240 µV, given that it is larger than 210 µV is `97.67%`

*(e) Estimate the interquartile range.*

We use the quantile function `qnorm` to find the first quartile (q1) and the third quartile (q3). Then, we calculate the interquartile range by subtracting q1 from q3.

```{r}
q1 <- qnorm(0.25, mean, std_dev)
q3 <- qnorm(0.75, mean, std_dev)
interqrange <- q3 - q1
interqrange
```

The estimated interquartile range is around `21.58`.

*(f) What is the probability that the signal will be less than 220 µV, given that it is larger than 210 µV?*

Similar to part (d), we calculate this probability by dividing the probability that the signal is between 210 and 220 µV by the probability that the signal is greater than 210 µV.

```{r}
probl220g210 <- (pnorm(220, mean, std_dev) - pnorm(210, mean, std_dev)) / (1 - pnorm(210, mean, std_dev))
probl220g210 * 100
```

The probability that the signal will be less than 220 µV, given that it is larger than 210 µV is `60.27%`

*(g) If we know that a received signal is greater that 200 µV, what is the probability that is in fact greater than 220 µV?*

We calculate this probability by dividing the probability that the signal is greater than 220 µV by the probability that the signal is greater than 200 µV.


```{r}
probg220g200 <- (1 - pnorm(220, mean, std_dev)) / (1 - pnorm(200, mean, std_dev))
probg220g200 * 100
```

The probability that it a received signal is in fact greater than 220 µV is `21.13%`

## 2. A manufacturer of a particular type of computer system is interested in improv- ing its customer support services. As a first step, its marketing department has been charged with the responsibility of summarizing the extent of customer problems in terms of system failures. Over a period of six months, customers were surveyed and the amount of downtime (in minutes) due to system failures they had experienced during the previous month was collected. The average downtime was found to be 25 minutes and a variance of 144. If it can be assumed that downtime is normally distributed:

First, we denote the average and variance downtime to 25 and 144, respectively.

```{r}
meandt <- 25
variancedt <- 144
```

Then we need to calculate the standard deviation because the normal distribution function requires it.Standard deviation is the square root of variance, so we use `sqrt()` function to find it.

```{r}
sddt <- sqrt(variancedt)
```

*(a) obtain bounds which will include 95% of the downtime of all the customers;*

To calculate bounds for 95% of the downtime, we use the `qnorm()` function. The qnorm() function calculates quantiles of the normal distribution.
We provide it with the probabilities 0.025 and 0.975 to get the lower and upper bounds respectively for 95% of the data.

```{r}
lowerbound95 <- qnorm(0.025, meandt, sddt)
upperbound95 <- qnorm(0.975, meandt, sddt)

lowerbound95
upperbound95
```

*(b) obtain the bound above which 10% of the downtime is included.*

Similarly, we use `qnorm()` function to find the quantile for 90% of the data. This gives us the bound above which 10% of the downtime is included.

```{r}
bound10 <- qnorm(0.90, meandt, sddt)
bound10
```
