---
title: "FA3 Probability"
author: "Lindsay Faith Bazar"
date: "February 26, 2024"
output: 
  cleanrmd::html_document_clean:
    theme: kacit
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**2.  A binary communication channel carries data as one of two sets of signals denoted by 0 and 1. Owing to noise, a transmitted 0 is sometimes received as a 1, and a transmitted 1 is sometimes received as a 0. For a given channel, it can be assumed that a transmitted 0 is correctly received with probability 0.95, and a transmitted 1 is correctly received with probability 0.75. Also, 70% of all messages are transmitted as a 0. If a signal is sent, determine the probability that:**

*(a) a 1 was received;*

To answer this given question, let us first denote the given probabilities:

The probability of receiving 0 given that 0 was transmitted is 0.95 denoted as `R0`:

```{r}
R0 <- 0.95
```

The 0.75 probability of receiving 1 given that 1 was transmitted and denoted as `R1`:

```{r}
R1 <- 0.75  

```

The 0.70 probability of transmitting 0 denoted as `T0`:

```{r}
T0 <- 0.70          
```

And lastly, the 0.30 probability of transmitting 1 denoted as `T1`:

```{r}
T1 <- 0.30          
```

We will denote the probability that a 1 was received as `R1Prob`. 

To find the probability, we can use the law of total probability:


```{r}
R1Prob <- (1 - R0) * T0 + R1 * T1
R1Prob
```

So, the probability that a 1 was received is `0.26`.

Now that we solved the probability of receiving 1, we can use Baye's theorem to find the probability of transmitting 1 when 1 is received.

*(b) a 1 was transmitted given than a 1 was received.*

We will denote this as `TR1`:

```{r}
TR1 <- R1 * T1 / R1Prob
TR1
```

So, the probability that a 1 was transmitted given that a 1 was received is approximately `0.865`.

**7.  There are three employees working at an IT company: Jane, Amy, and Ava, doing 10%, 30%, and 60% of the programming, respectively. 8% of Jane’s work, 5% of Amy’s work, and just 1% of Ava‘s work is in error.** 

*What is the overall percentage of error?* 

To find the overall percentage of error, let's first calculate the weighted error percentages for each employees based on their work contribution and percent of error.

`Jane`  
Error rate: 8%  
Work contribution: 10% 

```{r}
Jane <- .08 * .10 * 100
Jane
```


`Amy`  
Error rate: 5%  
Work Contribution: 30%

```{r}
Amy <- .05 * .30 * 100
Amy
```

`Ava`  
Error rate: 1%  
Work Contribution: 60%

```{r}
Ava <- .01 * .60 * 100
Ava
```

Now that we have solved the weighted error percentages for each of the employees, we can sum these values up to solve for the overall percentage of error.
```{r}
error <- Jane + Amy + Ava
error
```

So, the overall percentage of error is `2.9%`.

*If a program is found with an error, who is the most likely person to have written it?*

To find the person who most have likely written an error, we can compare each of the employees error percentages:

Jane: `0.8%`  
Amy: `1.5%`  
Ava: `0.6%`

Since Amy has the highest percentage of error (1.5%), `Amy` is the most likely person to have written a program with an error. 

